\documentclass[a4paper,11pt,twoside]{pads-thesis}

%%%%% PACKAGES USED SORTED ALPHABETICALLY  %%%%%
\usepackage[toc,page]{appendix} % support appendices
\usepackage{csquotes} % handles quotes automatically
\usepackage{hyperref} % allows autoref/hyperlinks/etc.
\usepackage{tikz}
\usepackage{tikzpeople}
\usepackage{tikzsymbols}
\usepackage{fontawesome}
\usepackage{figchild}
\usepackage{float}
\usepackage{stmaryrd}
\usepackage{fdsymbol}
\usepackage{tcolorbox}
\usepackage{colortbl}
\usepackage{esvect}



%%%%% PADS TEMPLATE SETTINGS %%%%%
\renewcommand{\gTitle}{Translucify: Enhancing Event Logs with Enabled Activities Beyond the Control-Flow Perspective \par}
\renewcommand{\gAuthor}{Geonho Yun}
\renewcommand{\gStudentID}{422305}
\renewcommand{\gHeader}{Bachelor's Thesis}
\renewcommand{\gSupervisor}{
    Harry H. Beyel, M.Sc. \\
    Christopher T. Schwanen, M.Sc. \\
    }

\renewcommand{\gExaminer}{Prof. Dr. Ir. Wil M. P. van der Aalst \\ 
Prof. Dr. rer. pol. Stefan Decker  }
\renewcommand{\gInstitution}{Chair of Process and Data Science (PADS - Informatik 9), RWTH Aachen University}
\renewcommand{\gRegDate}{2024-08-30}
\renewcommand{\gSubDate}{2024-10-21}

%%%%% AUTOREF %%%%%
\renewcommand{\sectionautorefname}{Section}

\begin{document}

%%% Title page
\pagenumbering{roman}

\gTitlePage
%\pagestyle{plain}
%%% TOC

\begin{abstract}
\chapter*{Abstract} \addcontentsline{toc}{chapter}{Abstract}
\input{chapters/abstract.tex}
\end{abstract}

\tableofcontents

%\setcounter{page}{1}
\cleardoublepage
%%% Begin document
\pagestyle{fancy}
\pagenumbering{arabic}

\chapter{Introduction} \label{chap:intro}
\input{chapters/introduction.tex}

\chapter{Related Work} \label{chap:related_work}
\input{chapters/related_work.tex}

\chapter{Preliminaries} \label{chap:prelim}
\input{chapters/prelim.tex}

\chapter{Translucifying the Event Log} \label{chap:method}
\input{chapters/method.tex}

\chapter{Implementation} \label{chap:impl}
\input{chapters/implementation.tex}

\chapter{Evaluation} \label{chap:eval}
\input{chapters/evaluation.tex}

\chapter{Discussion} \label{chap:discussion}
\input{chapters/discussion.tex}

\chapter{Conclusion} \label{chap:conclusion}
\input{chapters/conclusion.tex}

\bibliography{references} 
\addcontentsline{toc}{chapter}{Bibliography}

\begin{appendices}
    \chapter{Some Appendix}
    Try to avoid appendices.
    
\end{appendices}



\chapter*{Acknowledgments}\addcontentsline{toc}{chapter}{Acknowledgements}
\input{chapters/acknowledgements.tex}



\end{document}